### 2025-05-27 chapters/methodology.tex

\section{Experimental setup}

\section{Benchmarking tools and metrics}
%\section{Performance metrics}

\section{Parameter Variation Strategy}

\section{Data Collection and Analysis Procedures}