### 2025-06-06 main.tex

\documentclass[english, version-2020-11]{uzl-thesis}
\UzLStyle{pagella contrast design}
\usepackage{comment}
\usepackage{float}

\input{preamble.tex} % Put your \UzLStyle and \UzLThesisSetup here
\addbibresource{refs.bib}

\begin{document}

%\chapter{Introduction}

\include{chapters/introduction.tex}

\include{chapters/background.tex}

\include{chapters/methodology.tex}

\include{chapters/results.tex}

\include{chapters/recommendations.tex}

\include{chapters/conclusion.tex}

\end{document}