### 2025-06-16 chapters/discussion.tex
reorganised key sections

\chapter{Discussion}
\label{chap:discussion}

\section{Key Findings and Insights}

%\section{Challenges in Benchmarking Enclaves and Cryptographic Algorithms}

\section{Trade-offs Between Security and Performance in Enclaves}

\section{Implicaitons for Secure Computing and Cryptographic Applications}