### 2025-08-07 chapters/system-architecture.tex

\chapter{System Architecture and Design}
\label{chap:methodology}
The primary objective of this thesis is to assess the performance impact of Keystone’s enclave isolation mechanisms on typical embedded and systems-level workloads. Specifically, the study seeks to quantify the computational overhead introduced by executing applications within a Keystone enclave, as compared to their execution in a conventional, non-isolated environment. To achieve this, a series of carefully controlled experiments were designed and executed within a virtualized RISC-V system based on QEMU.

All benchmarks were executed in an emulated environment configured for the RV64GC architecture, which is representative of 64-bit RISC-V systems targeted by Keystone. The use of QEMU for emulation allowed for fine-grained control over the system environment, ensured repeatability, and eliminated the variability introduced by physical hardware, such as thermal throttling or hardware-specific optimizations. Although virtualized, the emulation accurately models instruction execution, privilege transitions, memory access patterns, and I/O behavior, making it suitable for preliminary performance evaluation.

To evaluate the impact of enclave execution, each benchmark was implemented as a standalone enclave application (referred to as an “eapp”), accompanied by a corresponding host application that manages enclave lifecycle events. The host application, running in user space, is responsible for initializing the enclave, loading the benchmark binary, invoking enclave entry points, and handling edge calls used for communication between the enclave and the host. The enclave application, in contrast, is isolated by the Keystone runtime and contains only the benchmark logic and internal data structures. This modular separation ensures that the same benchmark code can be executed both natively (without an enclave) and securely (within an enclave) without structural modification, enabling an accurate comparative analysis.

The evaluation involved executing each benchmark—Dhrystone and CoreMark—under two distinct configurations:

\begin{enumerate}
\item \textit{Native Execution:} The benchmark runs as a conventional user-space process within the QEMU-emulated Linux system, without invoking Keystone enclave services.
\item \textit{Enclave Execution:} The benchmark is loaded into a Keystone enclave, with isolation enforced by the Security Monitor using PMP.
\end{enumerate}

Each benchmark was run multiple times (typically ten iterations per configuration) to account for performance variability and ensure statistical robustness. For each run, metrics such as total execution time, throughput (measured in DMIPS for Dhrystone and iterations per second for CoreMark), and standard deviation were collected. The comparative analysis focused on the relative performance degradation observed in the enclave configuration, thereby providing a direct measurement of the cost of security.

This methodology enables a nuanced understanding of how Keystone’s isolation features influence real-world performance metrics. The controlled nature of the test environment, coupled with repeated measurement and the use of industry-standard benchmarks, ensures that the results are both meaningful and reproducible. By isolating the effect of enclave execution, this study contributes valuable insights into the trade-offs between security and efficiency in open-source TEE architectures.

\section{Experimental Setup}
\label{sec:experimental-setup}

The experimental setup was designed to enable reliable, repeatable evaluation of the Keystone Trusted Execution Environment (TEE) without requiring physical RISC-V hardware. To achieve this, a virtualized testbed was constructed using a RISC-V-specific fork of the QEMU emulator running on an Ubuntu 22.04 LTS host system. This emulated environment provides fine-grained control over system configuration, facilitating flexible testing of Keystone enclaves while accurately modeling a real RISC-V platform.

QEMU, an open-source hardware emulator, was configured to emulate the RV64GC (64-bit general-purpose) RISC-V architecture, fully compatible with Keystone’s software stack. This choice reflects Keystone’s target deployment on modern RISC-V platforms and allows access to enclave memory protection and privilege separation features managed by the Keystone Security Monitor. Ubuntu 22.04 LTS was selected for its long-term support, stability, and compatibility with the Keystone development toolchain, serving as a robust platform for building all components including the QEMU binary, Keystone kernel driver, runtime, and benchmark binaries.

The build process utilizes a combination of \texttt{Make} and Buildroot. Buildroot generates the embedded Linux root filesystem, cross-compiles necessary system libraries, and configures the Linux kernel running inside the QEMU guest. The Make-based system orchestrates compilation and deployment of Keystone components such as the runtime, Security Monitor, and kernel modules, ensuring reproducible builds targeting the QEMU RV64 environment.\footnote{Key configuration parameters include the QEMU platform, RV64 CPU architecture, and Buildroot settings optimized for Keystone enclave support.}

Launching the virtualized system is managed through a Makefile-driven workflow defined in the \texttt{run.mk} file, which encapsulates QEMU configuration and execution commands. Several runtime parameters are exposed as Makefile variables for flexibility, including:

\begin{itemize}
    \item \texttt{QEMU\_PORT} (default: 9821) — SSH port forwarding to access the QEMU guest.  
    \item \texttt{QEMU\_DBG\_PORT} (default: \texttt{QEMU\_PORT} + 1) — TCP port for GDB debugging.  
    \item \texttt{QEMU\_RAM\_SIZE} (default: 128M) — amount of emulated RAM.  
    \item \texttt{QEMU\_CORE\_COUNT} (default: 2) — number of virtual CPU cores.
\end{itemize}

QEMU is launched with flags specifying the machine type (\texttt{virt}), boot ROM and firmware, kernel and root filesystem images, VirtIO devices for block storage and networking, and network forwarding for SSH access. Optionally, a cache plugin \cite{mandour2021cache} can be enabled to provide detailed logging of cache events such as hits, misses, and evictions. This plugin offers deeper insight into the memory hierarchy’s behavior during enclave execution, aiding analysis of performance impacts due to cache and memory isolation overhead.

The cache plugin \cite{mandour2021cache} parameters are fully configurable at runtime, including:

\begin{itemize}
    \item Instruction cache: \texttt{icachesize}, \texttt{iblksize}, \texttt{iassoc} — size, block size, and associativity.  
    \item Data cache: \texttt{dcachesize}, \texttt{dblksize}, \texttt{dassoc} — analogous data cache parameters.  
    \item Eviction policy: \texttt{evict=lru|rand|fifo}.  
    \item Logging limits: \texttt{limit=TOP\_N} for number of top thrashing instructions to record.  
    \item Number of cores monitored: \texttt{cores=N\_CORES}.
\end{itemize}

For example, to configure the cache plugin with 8 KB L1 instruction and data caches, 64-byte blocks, 4-way associativity, 2 cores, and LRU eviction, the following command is used:

\begin{verbatim}
make QEMU_PLUGIN_ARGS="dcachesize=8192,dassoc=4,dblksize=64, \
icachesize=8192,iassoc=4,iblksize=64,cores=2,evict=lru" run
\end{verbatim}

While the plugin logs comprehensive cache activity, it cannot isolate cache events exclusively caused by enclave operations. Therefore, cache data interpretation requires caution when attributing performance effects specifically to enclaves.

Debugging support is available by enabling the \texttt{KEYSTONE\_DEBUG} flag, which starts QEMU’s built-in GDB server and halts execution at startup for remote debugging. After booting the QEMU guest, the Keystone kernel driver is loaded manually using:

\begin{verbatim}
modprobe keystone-driver
\end{verbatim}

This step initializes enclave functionality, allowing benchmark binaries to be securely loaded and executed within isolated memory regions managed by the Security Monitor.

Additional Makefile targets enable remote command execution over SSH and connection to the RISC-V GDB debugger for low-level inspection of enclave state. Debugging requires recompiling Keystone with debugging enabled, typically invoked as:

\begin{verbatim}
KEYSTONE_DEBUG=y make run
make debug-connect
\end{verbatim}

QEMU’s built-in GDB server provides visibility into registers, memory, and exceptions, aiding low-level inspection and troubleshooting of enclave operations.

Overall, this integrated virtualized environment and build system provide a flexible, controlled platform for detailed evaluation of Keystone’s performance and security properties. The modular Makefile-driven workflow streamlines configuration, deployment, and debugging, enabling efficient development and benchmarking without dependence on physical hardware. %Further technical details on build dependencies and configuration options are documented in the \textbf{Appendix}.

\section{Kyber in the Enclave}
\label{sec:kyber-enclave}

To comprehensively evaluate the performance characteristics of cryptographic workloads within a trusted execution environment, this study incorporates the Kyber key encapsulation mechanism (KEM), a lattice-based post-quantum cryptographic primitive. Kyber has garnered significant attention as a leading candidate in the NIST post-quantum cryptography standardization process, offering a robust balance of security, performance, and suitability for deployment on resource-constrained devices. Its integration within a secure enclave environment highlights the practical implications of deploying advanced cryptographic primitives where hardware-based isolation is critical for protecting sensitive operations and data.

The Kyber implementation employed in this work is derived from a well-maintained, lightweight C reference library, chosen for its portability and compatibility with embedded systems. This implementation was integrated into the Keystone enclave framework with minimal modification to preserve fidelity to the original cryptographic routines and avoid introducing artifacts that might bias performance measurement. Both the IND-CPA (indistinguishability under chosen-plaintext attack) and IND-CCA (indistinguishability under chosen-ciphertext attack) variants of Kyber were considered, with the primary cryptographic operations — key generation, encapsulation, and decapsulation — being individually invoked inside the enclave to facilitate granular performance profiling.

The following critical Kyber functions were separately instrumented and timed to capture their distinct computational demands:

\begin{itemize}
    \item \texttt{indcpa\_keypair} — Key pair generation for the IND-CPA variant.
    \item \texttt{indcpa\_enc} — Encryption (encapsulation) under IND-CPA assumptions.
    \item \texttt{indcpa\_dec} — Decryption (decapsulation) under IND-CPA assumptions.
    \item \texttt{kyber\_keypair} — Enhanced key pair generation supporting IND-CCA.
    \item \texttt{kyber\_encaps} — IND-CCA-compliant encapsulation.
    \item \texttt{kyber\_decaps} — IND-CCA-compliant decapsulation.
\end{itemize}

Each of these operations was executed across the five representative system configurations defined in Section~\ref{sec:param-variation}, which span a broad range of memory capacities, core counts, and execution modes. To ensure measurement accuracy and reduce variability from transient system states, ten iterations were run per configuration with the first iteration discarded as a warm-up phase. Execution timing was obtained using the enclave’s internal high-resolution timer API, capturing secure-world runtime exclusively. In parallel, external system-level metrics — such as CPU usage, memory footprint, and context switching counts — were gathered from the host environment to provide contextual insights into system overhead and resource utilization.

Unlike the synthetic benchmarks Dhrystone and CoreMark, Kyber’s workload is characterized by its intensive use of polynomial arithmetic, noise sampling, and number-theoretic transform (NTT) operations, all of which impose a distinct computational profile. Notably, Kyber exhibits a predominantly CPU-bound execution pattern, with relatively limited memory demands. This profile contrasts with cache- and memory-intensive workloads, which often demonstrate sensitivity to cache size, associativity, and eviction policies. Additionally, Kyber’s current implementation executes sequentially without exploiting parallelism or multithreading, making it well-suited for analysis within single-threaded enclave contexts.

It is important to note that the QEMU cache modeling plugin \cite{mandour2021cache} utilized for this study provides comprehensive logging of cache events and misses across the entire emulated system rather than isolating activity specific to the enclave or Kyber operations. Consequently, while cache plugin data enriches understanding of system-wide memory behavior, it does not enable direct attribution of cache hits or misses exclusively to Kyber’s execution. This limitation underscores the challenges of fine-grained performance attribution within complex emulation environments but does not diminish the value of combined timing and system-level metrics in capturing overall workload behavior.

Collectively, the measurement approach — leveraging both enclave-internal timers and host-level monitoring — facilitates a nuanced characterization of Kyber’s execution within a trusted environment. This methodology reveals how cryptographic routines with intricate arithmetic and memory access patterns perform under enclave constraints, including the overhead introduced by secure isolation and the impact of varying system resources such as memory size and CPU core availability.

Kyber’s inclusion in the benchmark suite extends the evaluation beyond traditional CPU-bound synthetic workloads, providing a practical example of a security-critical, compute-heavy application with significant implications for post-quantum readiness in embedded and secure computing contexts. The insights derived from profiling Kyber within Keystone enclaves inform considerations around enclave design, resource allocation, and the feasibility of deploying advanced cryptographic primitives in isolated environments with minimal performance degradation.

Kyber serves as a representative case study for secure, non-parallelizable cryptographic workloads executed within enclave boundaries. Its performance characteristics complement those observed in more generic benchmarks by illuminating the interplay between cryptographic computation, enclave isolation overhead, and system resource scaling. This comprehensive evaluation underpins recommendations for enclave provisioning and optimization in future secure computing platforms.

\section{Benchmarking tools and metrics}
\label{sec:benchmarking-tools}

To evaluate the performance impact of the Keystone Trusted Execution Environment (TEE) comprehensively, this study employs two well-established synthetic benchmarks: \textit{Dhrystone} and \textit{CoreMark}. Both are widely used in the embedded systems community and are well-supported in the RISC-V ecosystem, making them natural choices to assess Keystone’s behavior in secure versus non-secure execution modes. These benchmarks capture a broad spectrum of processor activities relevant to enclave execution—such as integer arithmetic, control flow, and memory access patterns.

The Dhrystone benchmark~\cite{weiss2002dhrystone} has been used for decades to measure general-purpose processor performance, especially in scenarios where floating-point operations and file I/O are not critical. Its focus on integer calculations and control structures—such as loops, function calls, and basic data handling—is representative of many embedded and real-time applications. Thanks to its simplicity and minimal memory usage, Dhrystone~\cite{weiss2002dhrystone} provides a clean baseline measure of raw processor capability. For this study, the RISC-V implementation of Dhrystone was obtained from the official \texttt{riscv-tests} repository to ensure accurate execution on the target platform. The benchmark was executed twice under identical conditions—once in native, non-secure mode and once inside a Keystone enclave. This dual setup facilitates precise measurement of the performance overhead imposed by Keystone’s security mechanisms, including memory isolation and context-switching between secure and non-secure worlds.

Dhrystone results are typically reported as "Dhrystones per second," quantifying how many benchmark iterations complete in one second. For broader comparability across different processor architectures, these results are normalized to Dhrystone MIPS (DMIPS), which estimates the effective instruction throughput in millions of instructions per second. This normalization allows meaningful performance comparisons by smoothing out architectural differences. In this study, comparing DMIPS between enclave and native executions highlights the raw performance impact of enabling the TEE.

To complement Dhrystone and gain insight into more complex system behavior, the CoreMark benchmark~\cite{gal2012exploring} was also employed. Developed by the Embedded Microprocessor Benchmark Consortium (EEMBC), CoreMark~\cite{gal2012exploring} addresses some limitations of Dhrystone by incorporating workloads that are closer to real-world embedded applications. It includes linked list manipulations, matrix operations, and finite state machine processing—tasks common in embedded software. Although still CPU-centric, CoreMark~\cite{gal2012exploring} exercises the memory subsystem more heavily through dynamic data structures, thereby exposing performance effects related to cache usage and memory protection features within a TEE.

The CoreMark version used in this study was sourced from the SiFive repository on GitHub\footnote{\url{https://github.com/sifive/benchmark-coremark}, accessed August 2025}, providing a RISC-V-optimized build compatible with Keystone. Like Dhrystone, CoreMark was run in both native and enclave modes, with results reported as "Iterations Per Second" (IPS), indicating how many complete benchmark runs occur per second. Because CoreMark combines arithmetic and memory operations, IPS delivers a more comprehensive view of system performance, reflecting both CPU and memory subsystem efficiency.

In addition to primary performance metrics—DMIPS for Dhrystone and IPS for CoreMark—this study also analyzed supporting metrics to better characterize system behavior. Overall execution time, measured in wall-clock seconds, was recorded for each benchmark run. Comparing execution times between enclave and native modes quantifies latency introduced by secure execution, including overheads from memory protection, context switching, and enclave isolation mechanisms.

Variability across runs was assessed through the standard deviation of benchmark results over multiple iterations. A low standard deviation signifies consistent and reliable performance, whereas higher variability might indicate sensitivity to system noise, scheduling delays, or other environmental factors.

Both Dhrystone~\cite{weiss2002dhrystone} and CoreMark~\cite{gal2012exploring} are CPU-intensive benchmarks, making them particularly useful for this study’s focus on processor and memory subsystem behavior under secure execution. Dhrystone targets raw integer and control flow performance, offering a clear view of base processor throughput, while CoreMark’s more sophisticated workload reveals broader impacts of the TEE’s security mechanisms, especially regarding memory interactions. Neither benchmark involves file I/O or persistent storage; therefore, this study does not assess storage-related TEE features such as secure disk access or file encryption, which would require alternative benchmarking approaches.

By combining these two benchmarks with detailed timing and variability analysis, this study provides a well-rounded and nuanced understanding of the trade-offs involved in securing execution via Keystone enclaves on RISC-V platforms.

\section{Parameter Variation Strategy}
\label{sec:param-variation}

To investigate the architectural impact on enclave execution, this study systematically varied system parameters including memory size, core count, execution mode (sequential vs parallel), and cache configurations using QEMU’s cache modeling capabilities. The aim was to understand how these factors influence the behavior of trusted and non-trusted workloads under different resource constraints.

Cache simulation was conducted using the QEMU TCG plugin framework \cite{mandour2021cache}, which models per-core instruction and data caches in an idealized fashion. While this plugin is not cycle-accurate and does not isolate enclave-specific behavior, it enables consistent modeling of cache accesses and miss patterns across the entire QEMU runtime. This was sufficient to observe cache-related effects caused by workload behavior in different configurations.

To avoid excessive experimental overhead and redundant trends, five representative system configurations were selected. These were chosen to cover a wide range of memory capacities, CPU core counts, and execution modes. Each configuration enables a specific type of insight, ensuring a diverse but manageable evaluation matrix.

\begin{table}[h]
\centering
\begin{tabular}{|l|c|c|c|p{7.5cm}|}
\hline
\textbf{Config Name} & \textbf{RAM} & \textbf{Cores} & \textbf{Execution} & \textbf{Rationale} \\
\hline
Low-End Baseline     & 64M         & 1             & Sequential         & Serves as the minimal configuration to establish baseline performance. Useful for understanding enclave overhead under tight resource constraints. \\
\hline
Balanced Parallel    & 128M        & 2             & Parallel           & Captures early-stage multicore scaling behavior. Highlights effects of light parallelism under modest memory availability. \\
\hline
Mid-Range Sequential & 256M        & 4             & Sequential         & Enables evaluation of scaling effects due to additional cores and memory, while keeping execution sequential to avoid parallel noise. \\
\hline
High-End Parallel    & 512M        & 6             & Parallel           & Represents a well-provisioned system where both memory and core resources are sufficient. Useful for observing near-ideal parallel execution. \\
\hline
Max Capacity Stress  & 2G          & 10            & Parallel           & Tests the upper bounds of simulated platform capacity. Helps identify performance saturation and the point of diminishing returns. \\
\hline
\end{tabular}
\caption{Summary of representative configurations used in performance evaluation.}
\end{table}

Several other configurations—such as single-core setups with excessive memory or high-core setups with limited memory—were explored during preliminary testing. However, they offered little additional insight beyond what was observable in the selected set. For instance, excessive memory on a single core often led to underutilization, while many-core systems with small memory pools frequently resulted in excessive context switching. Both behaviors were captured effectively by existing selected configurations.

Across all system configurations, three cache profiles were tested: a baseline setup with 8\,KB 4-way LRU caches, a larger 32\,KB 8-way configuration, and a fixed-size cache using FIFO and random eviction strategies. A consistent block size of 64 bytes was used to preserve spatial locality. Although the cache plugin logs system-wide activity, observed trends were analyzed in conjunction with benchmark behavior to infer their likely contributions to overall performance variation.

Benchmarks were run in both native and enclave execution modes to enable direct comparisons. This design facilitated the isolation of architectural performance trends and the specific overhead introduced by trusted execution.

\section{Data Collection and Analysis Procedures}
\label{sec:data-collection}

All benchmarks were executed within a consistent virtualized RISC-V environment using QEMU. For each selected configuration, both Dhrystone and CoreMark benchmarks were run in native and enclave modes to capture the impact of enclave isolation under different resource and cache conditions.

Each experiment consisted of ten executions. To reduce variance caused by warm-up behavior, the first run was discarded, and the remaining nine were used to compute average performance. Key performance metrics included total wall-clock execution time, benchmark throughput (DMIPS or iterations/sec), and standard deviation.

In addition to these primary measurements, system-level runtime statistics were collected from within the guest Linux environment. These included:

\begin{itemize}
    \item User and system CPU time
    \item Maximum memory usage (RSS)
    \item Minor and major page faults
    \item Voluntary and involuntary context switches
\end{itemize}

These metrics were essential to understanding how different system configurations affected workload behavior. For example, high numbers of involuntary context switches in parallel configurations often indicated memory contention or over-subscription of CPU cores.

Cache-related metrics were gathered using the QEMU cache plugin, which logs all instruction and data cache accesses and misses at the system level. Although enclave-specific cache statistics could not be isolated, the relative trends between native and enclave executions under identical configurations allowed for meaningful comparative analysis.

Performance overhead introduced by enclave execution was quantified using the percentage increase in execution time compared to native mode, calculated as:

\[
\text{Overhead (\%)} = \left( \frac{\text{Enclave Time} - \text{Native Time}}{\text{Native Time}} \right) \times 100
\]

This normalized metric enabled consistent cross-comparison between configurations with vastly different execution baselines.

All benchmark scripts, plugin configurations, and raw logs were version-controlled and archived. Detailed plugin output, configuration parameters, and system call traces are provided in Appendix~\ref{appendix:cache-plugin} for transparency and reproducibility.