### 2025-08-18 main.tex

\documentclass[english, version-2020-11]{uzl-thesis}
\UzLStyle{pagella contrast design}
\usepackage{comment}
\usepackage{fontspec}
\usepackage{float}
\lstset{basicstyle=\ttfamily}
\usepackage{tikz}
\usetikzlibrary{matrix}
\usepackage{pgfplots}
\usetikzlibrary{pgfplots.fillbetween}
\pgfplotsset{compat=1.18}
\usepackage{pgfplotstable}
\usepackage{pgf-umlsd}
\usepgfplotslibrary{statistics}
\usetikzlibrary{patterns} 
\usepackage{subcaption}
\usepackage{booktabs}
\usepackage{siunitx}
\usepackage[flushleft]{threeparttable}
\usepackage{enumitem}
\sloppy
\usetikzlibrary{shapes.geometric, arrows.meta, positioning, shadows.blur}
\usetikzlibrary{positioning,backgrounds,calc,shapes, shapes,arrows.meta}
\usepackage{xcolor}
\lstset{
    basicstyle=\ttfamily\small,
    frame=single,
    columns=fullflexible,
    breaklines=true,
    backgroundcolor=\color{gray!5},
    keywordstyle=\color{blue!70!black},
    commentstyle=\color{gray!60},
    stringstyle=\color{red!70!black},
    captionpos=b,
    showstringspaces=false,
    tabsize=4
}

\tikzstyle{process} = [
    rectangle, rounded corners,
    minimum width=4.2cm,
    minimum height=1.2cm,
    text centered,
    draw=black,
    fill=blue!10,
    font=\sffamily\small,
    blur shadow
]

\tikzstyle{arrow} = [
    thick,
    ->,
    >=Stealth,
    draw=gray!80
]

\tikzstyle{cell} = [rectangle, draw=black, minimum width=2.5em, minimum height=2.5em, anchor=center]
\tikzstyle{native} = [cell, fill=blue!20]
\tikzstyle{enclave} = [cell, fill=red!20]
\tikzstyle{label} = [font=\bfseries]

\input{preamble.tex} % Put your \UzLStyle and \UzLThesisSetup here
\addbibresource{refs.bib}

\begin{document}

%\chapter{Introduction}

\include{chapters/introduction.tex}

\include{chapters/background.tex}

\include{chapters/system-architecture.tex}

\include{chapters/benchmark.tex}

%\include{chapters/results.tex}

\include{chapters/discussion.tex}

%\include{chapters/recommendations.tex}

\include{chapters/conclusion.tex}

\end{document}