### 2025-08-23 chapters/system-architecture.tex
chapter 3 complete

\chapter{System Design and Methodology}
\label{chap:methodology}

The primary objective of this thesis is to evaluate the performance impact of Keystone’s enclave isolation mechanisms on representative embedded and systems-level workloads. Specifically, the study aims to quantify the computational overhead introduced by executing applications within a Keystone enclave, compared to their execution in a traditional, non-isolated environment. To this end, a series of controlled experiments were conducted on a virtualized RISC-V platform based on QEMU.

All benchmarks were executed within an emulated environment configured for the RV64GC architecture, which corresponds to the 64-bit RISC-V ISA targeted by Keystone. The QEMU-based emulation provides fine-grained control over system parameters and ensures experimental repeatability by eliminating variability introduced by physical hardware—such as thermal throttling, cache behavior, or platform-specific optimizations. While virtualized, QEMU accurately models instruction execution, privilege-level transitions, memory access patterns, and I/O behavior, making it a viable platform for preliminary performance evaluations of trusted execution environments.

To assess the performance implications of enclave-based isolation, each benchmark was implemented as a standalone enclave application (referred to as an \textit{eapp}), accompanied by a corresponding host application responsible for enclave lifecycle management. The host application, running in user space, handles enclave initialization, benchmark loading, invocation of enclave entry points, and inter-domain communication via edge calls. In contrast, the enclave application is fully isolated by the Keystone runtime and contains only the benchmark logic and associated data structures. This strict separation ensures that the same benchmark codebase can be executed both natively (without isolation) and securely (within an enclave), without requiring structural changes—thus enabling a fair and consistent comparative analysis.

The experimental procedure followed a systematic pipeline, outlined in Figure~\ref{fig:benchmarking-swimlane}, beginning with benchmark selection and system configuration, followed by dual-mode execution (native and enclave), and concluding with metric collection and comparative analysis.

\begin{figure}[H]
\centering
\begin{tikzpicture}[
    >=latex,
    node distance=1.4cm and 3.5cm,
    lane/.style={
        rectangle,
        draw=black!80,
        thick,
        minimum width=6.5cm,
        minimum height=0.8cm,
        rounded corners,
        align=center,
        text width=6.2cm,
        fill=gray!5
    },
    arrow/.style={->, very thick},
    laneTitle/.style={
        font=\bfseries,
        fill=blue!15,
        rounded corners,
        draw=black!80,
        minimum height=0.8cm
    }
]

% Lane titles
\node[laneTitle, text width=6.5cm] (lane1) {Setup \& Configuration};
\node[laneTitle, right=of lane1]   (lane2) {Execution \& Analysis};

% Setup lane nodes
\node[lane, below=of lane1] (start)   {\textbf{Define benchmark set}};
\node[lane, below=of start] (config)  {\textbf{Select hardware configuration}};

% Execution lane nodes
\node[lane, below=of lane2] (native)  {\textbf{Run benchmarks in native mode}};
\node[lane, below=of native] (enclave) {\textbf{Run benchmarks in enclave mode}};
\node[lane, below=of enclave] (metrics) {\textbf{Collect performance metrics}};
\node[lane, below=of metrics] (analyze) {\textbf{Compare and analyze overheads}};
\node[lane, fill=green!10, below=of analyze] (end) {\textbf{Store logs and archive results}};

% Arrows
\draw[arrow] (start) -- (config);
\draw[arrow] (config.east) -- ++(0.8,0) -- ++(0,-0.7) -- (native.west);
\draw[arrow] (native) -- (enclave);
\draw[arrow] (enclave) -- (metrics);
\draw[arrow] (metrics) -- (analyze);
\draw[arrow] (analyze) -- (end);

\end{tikzpicture}
\caption{Benchmarking and data collection workflow}
\label{fig:benchmarking-swimlane}
\end{figure}

The experimental evaluation was conducted using two well-established benchmarking tools—Dhrystone and CoreMark—each executed under two configurations:

\begin{enumerate}
\item \textbf{Native Execution:} The benchmark runs as a conventional user-space process within the QEMU-emulated Linux system, with no involvement of Keystone's enclave infrastructure.
\item \textbf{Enclave Execution:} The benchmark is loaded into a Keystone enclave, where memory isolation is enforced by the Security Monitor through RISC-V Physical Memory Protection (PMP) mechanisms.
\end{enumerate}

Each benchmark configuration was executed multiple times (typically ten iterations) to mitigate performance variance and ensure statistical robustness. For each run, performance metrics including total execution time, throughput (expressed in DMIPS for Dhrystone and iterations per second for CoreMark), and standard deviation were collected. The analysis focuses on the relative performance degradation incurred in the enclave execution scenario, thereby providing a direct estimate of the overhead associated with enclave-based isolation.

This methodology facilitates a detailed understanding of how Keystone’s architectural design and isolation primitives influence runtime performance. The controlled and repeatable test environment, combined with standardized benchmarking tools and consistent measurement practices, ensures that the results are both reliable and meaningful. Ultimately, the findings offer valuable insights into the trade-offs between security and efficiency in open-source TEE frameworks and inform future efforts in trusted system design.

\section{Experimental Platform and Environment}
\label{sec:experimental-setup}

The experimental setup was designed to enable reliable, repeatable evaluation of the Keystone Trusted Execution Environment (TEE) without requiring physical RISC-V hardware. To achieve this, a virtualized testbed was constructed using a RISC-V-specific fork of the QEMU emulator running on an Ubuntu 22.04 LTS host system. This emulated environment provides fine-grained control over system configuration, facilitating flexible testing of Keystone enclaves while accurately modeling a real RISC-V platform.

QEMU, an open-source hardware emulator, was configured to emulate the RV64GC (64-bit general-purpose) RISC-V architecture, fully compatible with Keystone’s software stack. This choice reflects Keystone’s target deployment on modern RISC-V platforms and allows access to enclave memory protection and privilege separation features managed by the Keystone Security Monitor. Ubuntu 22.04 LTS was selected for its long-term support, stability, and compatibility with the Keystone development toolchain, serving as a robust platform for building all components including the QEMU binary, Keystone kernel driver, runtime, and benchmark binaries.

Figure~\ref{fig:integrated_two_column_peripherals_below} provides a unified representation of the development workflow and the virtualized execution environment used in this thesis. On the left, the build process begins with the Keystone source code and proceeds through root filesystem generation via Buildroot, orchestration of key components through a \texttt{Make}-based system, and culminates in the launch of the full system within a QEMU-emulated RV64GC RISC-V platform. On the right, the runtime stack is depicted in detail: Ubuntu 22.04 LTS acts as the host operating system, supporting QEMU, which emulates the Keystone stack—including the Security Monitor, runtime, and enclave applications.
\begin{figure}[htbp]
\centering
\begin{tikzpicture}[
    font=\scriptsize,
    node distance=0.8cm and 1.0cm,
    box/.style={
        draw,
        rounded corners,
        fill=gray!15,
        minimum width=3.6cm,
        minimum height=0.9cm,
        align=center
    },
    stack/.style={
        draw,
        rounded corners,
        fill=gray!15,
        minimum width=3.0cm,
        minimum height=0.9cm,
        align=center
    },
    keystone/.style={
        fill=white,
        thick,
        minimum width=3.0cm
    },
    arrow/.style={
        ->,
        thick,
        >=stealth
    }
]

% --- Left column: Build & Deployment Workflow ---
\node (src) [box] {Source Code \\ Keystone, Kernel, Apps};
\node (buildroot) [box, below=of src] {Buildroot \\ Rootfs + Kernel};
\node (make) [box, below=of buildroot] {Make Orchestration};
\node (launch) [box, below=of make] {QEMU Launch};
\node (bench) [box, below=of launch] {Run Enclave / Benchmark};

\foreach \from/\to in {src/buildroot, buildroot/make, make/launch, launch/bench} {
    \draw[arrow] (\from) -- (\to);
}

% --- Right column: Testbed Stack ---
\node (host) [stack, fill=gray!30, minimum height=1.1cm, right=5.0cm of src] {Ubuntu 22.04 LTS\\ Host OS};
\node (qemu) [stack, below=of host] {QEMU RV64GC Emulator};
\node (keystone) [stack, keystone, below=of qemu] {Keystone Stack\\ \footnotesize Sec. Monitor, Runtime, Enclaves};

% Peripherals BELOW Keystone, side-by-side
\node (virtio) [stack, fill=gray!20, below left=0.5cm and 0.5cm of keystone.south] {VirtIO Devices};
\node (plugin) [stack, fill=gray!20, below right=0.5cm and 0.5cm of keystone.south] {Cache Plugin};

% Testbed arrows
\draw[arrow] (host) -- (qemu);
\draw[arrow] (qemu) -- (keystone);
\draw[arrow] (keystone.south) -- ++(0,-0.25) -| (virtio.north);
\draw[arrow] (keystone.south) -- ++(0,-0.25) -| (plugin.north);

% Cross-links from workflow to stack (no intersection)
\draw[arrow, dashed] (launch.east) -- ++(0.8,0) |- (qemu.west);
\draw[arrow, dashed] (bench.east)  -- ++(1.2,0) |- (keystone.west);

% Annotation
\node[font=\tiny, anchor=south west] at ([xshift=0.15cm,yshift=0.15cm]host.north west) {SSH/GDB access};

\end{tikzpicture}
\caption{Overview of the build workflow and virtualized Keystone testbed. The left column shows the step-by-step process of building and launching enclave applications, starting from source code to benchmark execution. The right column depicts the runtime testbed architecture on Ubuntu using QEMU for RISC-V emulation, including the Keystone software stack and supporting peripheral modules (VirtIO and cache plugin). Dashed arrows indicate execution-stage linkage between build steps and their corresponding runtime components.}
\label{fig:integrated_two_column_peripherals_below}
\end{figure}

The design supports modular and reproducible deployment. Peripheral components such as VirtIO devices and an optional cache plugin are attached beneath the enclave runtime, enabling I/O functionality and microarchitectural performance instrumentation. Dashed arrows visually link specific build stages to their corresponding runtime targets, providing a clear and traceable path from compilation to execution. This integration highlights how the build system and execution environment are tightly coupled to support secure enclave development and testing.\footnote{Key configuration parameters include the QEMU platform target, RV64GC architecture, and Buildroot settings customized for Keystone compatibility.}

Launching the system is handled via a Makefile-based interface, which wraps QEMU commands with tunable runtime parameters. The most relevant configuration variables and their default values are summarized in Table~\ref{tab:qemu-params}, including RAM size, core count, and SSH port mapping.
\begin{table}[h]
\centering
\caption{Key QEMU launch parameters and defaults.}
\label{tab:qemu-params}
\begin{tabular}{l l p{6cm}}
\toprule
\textbf{Parameter} & \textbf{Default} & \textbf{Description} \\
\midrule
QEMU\_PORT & 9821 & SSH port forwarding to guest \\
QEMU\_DBG\_PORT & QEMU\_PORT+1 & TCP port for GDB debugging \\
QEMU\_RAM\_SIZE & 128M & Emulated RAM size \\
QEMU\_CORE\_COUNT & 2 & Number of virtual CPU cores \\
\bottomrule
\end{tabular}
\end{table}

Optionally, the cache plugin~\cite{mandour2021cache} can be integrated into the QEMU runtime. This plugin enables logging of cache events—such as hits, misses, and evictions—and is particularly useful when analyzing memory performance of the overall system. Its configuration options are shown in Listing~\ref{lst:cache_config}.
%\vspace{5em}
\begin{lstlisting}[language=bash,caption={Configuration options for cache simulation},label={lst:cache_config}]
# Instruction cache parameters
icachesize=SIZE         # Size of instruction cache
iblksize=BLOCK_SIZE     # Block size
iassoc=ASSOCIATIVITY    # Associativity

# Data cache parameters
dcachesize=SIZE         # Size of data cache
dblksize=BLOCK_SIZE     # Block size
dassoc=ASSOCIATIVITY    # Associativity

# Cache eviction policy
evict=lru|rand|fifo     # Replacement policy

# Logging configuration
limit=TOP_N             # Number of top thrashing instructions to record

# Monitoring settings
cores=N_CORES           # Number of cores to monitor
\end{lstlisting}

For example, Listing~\ref{lst:cache_cmd} demonstrates how to configure an L1 cache setup at runtime.

\begin{lstlisting}[language=bash, caption={Configuration command for cache plugin}, label={lst:cache_cmd}]
make QEMU_PLUGIN_ARGS="dcachesize=8192,dassoc=4,dblksize=64, \
icachesize=8192,iassoc=4,iblksize=64,cores=2,evict=lru" run
\end{lstlisting}

While the plugin logs comprehensive cache activity, it cannot isolate cache events exclusively caused by enclave operations. Therefore, cache data interpretation requires caution when attributing performance effects specifically to enclaves.

For kernel-level enclave debugging, QEMU’s built-in GDB server can be activated via the \texttt{KEYSTONE\_DEBUG} flag. Once the system boots, the Keystone kernel driver is loaded manually using the command in Listing~\ref{lst:modprobe}.

\begin{lstlisting}[language=bash, caption={Loading the Keystone kernel driver}, label={lst:modprobe}]
modprobe keystone-driver
\end{lstlisting}

This step initializes enclave functionality, allowing benchmark binaries to be securely loaded and executed within isolated memory regions managed by the Security Monitor.

In debugging scenarios, the Makefile target \texttt{debug-connect} can be used to attach a GDB session to the running guest. A full debugging session typically starts as shown in Listing~\ref{lst:debug}.

\begin{lstlisting}[language=bash, label={lst:debug}, caption={Starting a QEMU session with GDB debugging enabled}]
KEYSTONE_DEBUG=y make run
make debug-connect
\end{lstlisting}

QEMU’s built-in GDB server provides visibility into registers, memory, and exceptions, aiding low-level inspection and troubleshooting of enclave operations.

Overall, this integrated virtualized environment and build system provide a flexible, controlled platform for detailed evaluation of Keystone’s performance and security properties. The modular Makefile-driven workflow streamlines configuration, deployment, and debugging, enabling efficient development and benchmarking without dependence on physical hardware. %Further technical details on build dependencies and configuration options are documented in the \textbf{Appendix}.

\section{Integration of Kyber into Keystone}
\label{sec:kyber-enclave}

To comprehensively evaluate the performance characteristics of cryptographic workloads within a trusted execution environment, this study incorporates the Kyber key encapsulation mechanism (KEM), a lattice-based post-quantum cryptographic primitive. Kyber has garnered significant attention as a leading candidate in the NIST post-quantum cryptography standardization process, offering a robust balance of security, performance, and suitability for deployment on resource-constrained devices. Its integration within a secure enclave environment highlights the practical implications of deploying advanced cryptographic primitives where hardware-based isolation is critical for protecting sensitive operations and data.

The Kyber implementation employed in this work is derived from a lightweight, standards-compliant C reference library, selected for its portability and compatibility with embedded systems. This implementation was integrated into the Keystone enclave framework with minimal modification to preserve the correctness and performance characteristics of the original routines, while avoiding structural changes that might distort measurement fidelity.

Both the IND-CPA (indistinguishability under chosen-plaintext attack) and IND-CCA (indistinguishability under chosen-ciphertext attack) security models were considered. For performance profiling, the core cryptographic routines—key generation, encapsulation, and decapsulation—were individually invoked within the enclave. This allowed for fine-grained instrumentation and measurement of each function’s runtime behavior under enclave-induced constraints.

The specific functions evaluated are shown in Listing~\ref{lst:kyber_api}, grouped by security level:

\begin{lstlisting}[language=bash,caption={Kyber cryptographic API functions},label={lst:kyber_api}]
indcpa_keypair   # Key pair generation for IND-CPA variant
indcpa_enc       # Encryption (encapsulation) under IND-CPA assumptions
indcpa_dec       # Decryption (decapsulation) under IND-CPA assumptions

kyber_keypair    # Enhanced key pair generation supporting IND-CCA
kyber_encaps     # IND-CCA-compliant encapsulation
kyber_decaps     # IND-CCA-compliant decapsulation
\end{lstlisting}
Each of these operations was executed across the five representative system configurations defined in Section~\ref{sec:param-variation}, which span a broad range of memory capacities, core counts, and execution modes. To ensure measurement accuracy and reduce variability from transient system states, ten iterations were run per configuration with the first iteration discarded as a warm-up phase. Execution timing was obtained using the enclave’s internal high-resolution timer API, capturing secure-world runtime exclusively. In parallel, external system-level metrics — such as CPU usage, memory footprint, and context switching counts — were gathered from the host environment to provide contextual insights into system overhead and resource utilization.

Unlike the synthetic benchmarks Dhrystone and CoreMark, Kyber’s workload is characterized by its intensive use of polynomial arithmetic, noise sampling, and number-theoretic transform (NTT) operations, all of which impose a distinct computational profile. Notably, Kyber exhibits a predominantly CPU-bound execution pattern, with relatively limited memory demands. This profile contrasts with cache- and memory-intensive workloads, which often demonstrate sensitivity to cache size, associativity, and eviction policies. Additionally, Kyber’s current implementation executes sequentially without exploiting parallelism or multithreading, making it well-suited for analysis within single-threaded enclave contexts.

It is important to note that the QEMU cache modeling plugin \cite{mandour2021cache} utilized for this study provides comprehensive logging of cache events and misses across the entire emulated system rather than isolating activity specific to the enclave or Kyber operations. Consequently, while cache plugin data enriches understanding of system-wide memory behavior, it does not enable direct attribution of cache hits or misses exclusively to Kyber’s execution. This limitation underscores the challenges of fine-grained performance attribution within complex emulation environments but does not diminish the value of combined timing and system-level metrics in capturing overall workload behavior.

Collectively, the measurement approach — leveraging both enclave-internal timers and host-level monitoring — facilitates a nuanced characterization of Kyber’s execution within a trusted environment. This methodology reveals how cryptographic routines with intricate arithmetic and memory access patterns perform under enclave constraints, including the overhead introduced by secure isolation and the impact of varying system resources such as memory size and CPU core availability.

Kyber’s inclusion in the benchmark suite extends the evaluation beyond traditional CPU-bound synthetic workloads, providing a practical example of a security-critical, compute-heavy application with significant implications for post-quantum readiness in embedded and secure computing contexts. The insights derived from profiling Kyber within Keystone enclaves inform considerations around enclave design, resource allocation, and the feasibility of deploying advanced cryptographic primitives in isolated environments with minimal performance degradation.

Kyber serves as a representative case study for secure, non-parallelizable cryptographic workloads executed within enclave boundaries. Its performance characteristics complement those observed in more generic benchmarks by illuminating the interplay between cryptographic computation, enclave isolation overhead, and system resource scaling. This comprehensive evaluation underpins recommendations for enclave provisioning and optimization in future secure computing platforms.

\section{Benchmarking Suite and Metrics}
\label{sec:benchmarking-tools}

To evaluate the performance impact of the Keystone Trusted Execution Environment (TEE) comprehensively, this study employs two well-established synthetic benchmarks: \textit{Dhrystone} and \textit{CoreMark}. Both are widely used in the embedded systems community and are well-supported in the RISC-V ecosystem, making them natural choices to assess Keystone’s behavior in secure versus non-secure execution modes. These benchmarks capture a broad spectrum of processor activities relevant to enclave execution—such as integer arithmetic, control flow, and memory access patterns.

The Dhrystone benchmark~\cite{weiss2002dhrystone} has been used for decades to measure general-purpose processor performance, especially in scenarios where floating-point operations and file I/O are not critical. Its focus on integer calculations and control structures—such as loops, function calls, and basic data handling—is representative of many embedded and real-time applications. Thanks to its simplicity and minimal memory usage, Dhrystone~\cite{weiss2002dhrystone} provides a clean baseline measure of raw processor capability. For this study, the RISC-V implementation of Dhrystone was obtained from the official \texttt{riscv-tests} repository to ensure accurate execution on the target platform. The benchmark was executed twice under identical conditions—once in native, non-secure mode and once inside a Keystone enclave. This dual setup facilitates precise measurement of the performance overhead imposed by Keystone’s security mechanisms, including memory isolation and context-switching between secure and non-secure worlds.

Dhrystone results are typically reported as "Dhrystones per second," quantifying how many benchmark iterations complete in one second. For broader comparability across different processor architectures, these results are normalized to Dhrystone MIPS (DMIPS), which estimates the effective instruction throughput in millions of instructions per second. This normalization allows meaningful performance comparisons by smoothing out architectural differences. In this study, comparing DMIPS between enclave and native executions highlights the raw performance impact of enabling the TEE.

To complement Dhrystone and gain insight into more complex system behavior, the CoreMark benchmark~\cite{gal2012exploring} was also employed. Developed by the Embedded Microprocessor Benchmark Consortium (EEMBC), CoreMark~\cite{gal2012exploring} addresses some limitations of Dhrystone by incorporating workloads that are closer to real-world embedded applications. It includes linked list manipulations, matrix operations, and finite state machine processing—tasks common in embedded software. Although still CPU-centric, CoreMark~\cite{gal2012exploring} exercises the memory subsystem more heavily through dynamic data structures, thereby exposing performance effects related to cache usage and memory protection features within a TEE.

The CoreMark version used in this study was sourced from the SiFive repository on GitHub\footnote{\url{https://github.com/sifive/benchmark-coremark}, accessed August 2025}, providing a RISC-V-optimized build compatible with Keystone. Like Dhrystone, CoreMark was run in both native and enclave modes, with results reported as "Iterations Per Second" (IPS), indicating how many complete benchmark runs occur per second. Because CoreMark combines arithmetic and memory operations, IPS delivers a more comprehensive view of system performance, reflecting both CPU and memory subsystem efficiency.

As shown in Listing~\ref{lst:coremark-output}, the CoreMark benchmark outputs the number of iterations per second (IPS), which is a key metric for evaluating system performance.
\begin{lstlisting}[language=bash, caption={CoreMark sample output used to extract IPS}, label={lst:coremark-output}]
2K performance run parameters for coremark.
CoreMark Size    : 666
Total ticks      : 112738928
Iterations/Sec   : 937.43
Iterations       : 100000
\end{lstlisting}

In addition to primary performance metrics—DMIPS for Dhrystone and IPS for CoreMark—this study also analyzed supporting metrics to better characterize system behavior. Overall execution time, measured in wall-clock seconds, was recorded for each benchmark run. Comparing execution times between enclave and native modes quantifies latency introduced by secure execution, including overheads from memory protection, context switching, and enclave isolation mechanisms.

Variability across runs was assessed through the standard deviation of benchmark results over multiple iterations. A low standard deviation signifies consistent and reliable performance, whereas higher variability might indicate sensitivity to system noise, scheduling delays, or other environmental factors.

Both Dhrystone~\cite{weiss2002dhrystone} and CoreMark~\cite{gal2012exploring} are CPU-intensive benchmarks, making them particularly useful for this study’s focus on processor and memory subsystem behavior under secure execution. Dhrystone targets raw integer and control flow performance, offering a clear view of base processor throughput, while CoreMark’s more sophisticated workload reveals broader impacts of the TEE’s security mechanisms, especially regarding memory interactions. Neither benchmark involves file I/O or persistent storage; therefore, this study does not assess storage-related TEE features such as secure disk access or file encryption, which would require alternative benchmarking approaches.

By combining these two benchmarks with detailed timing and variability analysis, this study provides a well-rounded and nuanced understanding of the trade-offs involved in securing execution via Keystone enclaves on RISC-V platforms.

\section{Parameter Configuration and Variation}
\label{sec:param-variation}

To rigorously assess how architectural and system-level parameters influence the performance of workloads executed within secure enclaves, this study undertook a systematic exploration of configurable runtime characteristics on the simulated RISC-V platform. Specifically, we varied four principal dimensions: physical memory size, the number of processor cores, execution mode (sequential versus parallel), and cache configuration. These parameters were chosen based on their known relevance to workload performance and enclave behavior, and were adjusted within the bounds of what is supported by the QEMU-based virtual platform and the Keystone framework.

The overarching objective of this variation was to uncover the extent to which system-level constraints and resource allocations impact both trusted (enclave) and non-trusted (native) execution. By exploring diverse configurations, we aimed to characterize how resource limitations (e.g., constrained memory or limited cores) manifest in performance degradation, and to what degree secure execution amplifies or mitigates these effects.

To simulate cache behavior, we employed the QEMU TCG plugin framework, specifically leveraging the cache modeling capabilities introduced by Mandour et al. \cite{mandour2021cache}. This plugin facilitates the modeling of per-core instruction and data caches and enables the collection of aggregate system-wide cache statistics. While it does not provide cycle-accurate timing nor isolate enclave-specific cache behavior, its ability to log cache access and miss patterns uniformly across all benchmark executions proved sufficient for analyzing cache-related effects in both secure and non-secure modes. The insights gleaned from this analysis, though approximate, were instrumental in identifying general cache utilization trends under various workloads and resource configurations.

Given the expansive parameter space created by the combination of memory size, core count, execution mode, and cache settings, it was neither practical nor necessary to exhaustively benchmark every possible configuration. Instead, we curated a set of five representative system configurations that collectively span a meaningful range of the performance spectrum—from minimal, resource-constrained environments to well-provisioned, multicore systems operating at near full capacity. These configurations are detailed in Table~\ref{tab:configurations}.

\begin{table}[h]
\centering
\caption{Summary of representative configurations used in performance evaluation.}
\label{tab:configurations}
\footnotesize
\setlength{\tabcolsep}{5pt}
\renewcommand{\arraystretch}{1.2}
\begin{tabular}{lccccccl}
\toprule
\textbf{Configuration} & \textbf{RAM} & \textbf{Cores} & \textbf{Exec.} &
\textbf{D‑Cache} & \textbf{I‑Cache} & \textbf{Policy} & \textbf{Assoc./Blk (D/I)} \\
\midrule
Low‑End Baseline &
64\,MB & 1 & Seq. &
4\,KB & 4\,KB & LRU & 2‑way / 32B,\ 2‑way / 32B \\
%\hline
Balanced Parallel &
128\,MB & 2 & Par. &
8\,KB & 8\,KB & FIFO & 4‑way / 64B,\ 4‑way / 64B \\
%\hline
Mid‑Range Sequential &
256\,MB & 4 & Seq. &
16\,KB & 8\,KB & Random & 8‑way / 32B,\ 4‑way / 64B \\
%\hline
High‑End Parallel &
512\,MB & 6 & Par. &
8\,KB & 16\,KB & LRU & 2‑way / 64B,\ 8‑way / 32B \\
%\hline
Max Capacity Stress &
2\,GB & 10 & Par. &
32\,KB & 32\,KB & FIFO & 16‑way / 64B,\ 16‑way / 64B \\
\bottomrule
\end{tabular}
\end{table}

In preliminary testing, a wider range of system specifications was explored to ensure that the final selection of five configurations (Table~\ref{tab:configurations}) provided both diversity and coverage. Extreme cases—such as single‑core systems with disproportionately large memory or many‑core systems with severely limited memory—were excluded, as they tended to either underutilize resources or exhibit pathological behaviour (e.g., excessive paging or context‑switching) without yielding additional insight. For example, adding memory to a single‑core system had negligible impact due to CPU‑bound bottlenecks, whereas increasing core counts with insufficient memory frequently caused scheduler overhead and contention.

In addition to varying memory capacity, core count, and execution mode (sequential or parallel), each configuration was paired with a distinct cache arrangement to assess the sensitivity of enclave and native workloads to cache architecture. The cache parameters are summarised in Table~\ref{tab:configurations} and Listing~\ref{lst:cache_profiles}.

These profiles were designed to represent common cache configurations found in embedded and general-purpose processors, ranging from small, low-associativity caches to larger, highly associative designs. The eviction policies (LRU, FIFO, Random) were selected to explore how different replacement strategies affect performance under varying workload characteristics.

\begin{lstlisting}[language=bash,caption={Cache configuration profiles},label={lst:cache_profiles}]
# Small caches
Size: 4-8 KB
Associativity: Low
Eviction Policy: LRU

# Mid-sized caches
Size: 8-16 KB
Associativity: Mixed
Eviction Policy: FIFO or Random

# Large caches
Size: 32 KB
Associativity: High
Eviction Policy: FIFO
\end{lstlisting}

All caches used fixed block sizes of 32 or 64\,B, as listed in the table, to maintain consistent spatial locality across workloads. While the cache modelling plugin measures aggregate system statistics without distinguishing enclave from non‑enclave activity, observed trends in miss rates and access patterns were correlated with benchmark behaviour to help attribute performance differences.

For each of the five configurations, all benchmarks were executed in both native and enclave modes. This mirrored arrangement facilitated a clear isolation of enclave‑specific overheads from general architectural effects.

\section{Data Collection and Reproducibility}
\label{sec:data-collection}

To ensure the integrity, consistency, and scientific validity of the performance evaluation, all benchmark experiments were conducted within a tightly controlled and reproducible virtualized environment. The entire measurement campaign was executed on a simulated RISC-V platform using the QEMU emulator, which served as the foundation for both native and enclave-based workloads. This consistent virtualization layer eliminated hardware variability and allowed for precise control over system resources such as memory, core count, and cache characteristics.

Each system configuration, as defined in Section~\ref{sec:param-variation}, was evaluated using the Dhrystone, CoreMark, and Kyber benchmarks. For each benchmark, tests were run in both native (non-enclave) and enclave execution modes to facilitate direct comparative analysis of performance overheads introduced by trusted execution.

To reduce transient measurement noise and improve the robustness of collected data, each benchmark experiment was repeated ten times per configuration. The first execution was discarded in all cases to account for initialization artifacts such as dynamic binary translation warm-up and memory allocation caching within the QEMU runtime. The remaining nine runs were averaged to produce stable and representative performance figures. In addition to computing the arithmetic mean, the standard deviation was recorded to provide a measure of variability across runs.

The primary performance metric collected was total wall-clock execution time, measured from within the guest Linux environment. For the Kyber benchmark, high-resolution timing was implemented explicitly using the \texttt{clock\_gettime()} function, which was integrated into the benchmark source code prior to compilation. In contrast, both Dhrystone and CoreMark include built-in timing and throughput reporting mechanisms as part of their standard distributions. Accordingly, benchmark-specific performance metrics were used where applicable: Dhrystone MIPS (DMIPS) for Dhrystone, and iterations per second for CoreMark. These metrics enabled more nuanced comparisons of workload efficiency across various enclave and system configurations.

In addition to these high-level metrics, detailed system-level runtime statistics were gathered using standard Linux facilities such as the \texttt{getrusage()} system call.

As shown in Listing~\ref{lst:getrusage}, this call provides insights into CPU time, memory usage, page faults, and context switching events. Collecting these indicators was crucial for diagnosing performance behaviors that could not be explained solely by execution time or throughput.

% \begin{lstlisting}[language=bash]
% # CPU time: Total time spent in user mode and kernel mode.
% # Peak RSS: Maximum resident memory usage during execution.
% # Page faults: Number of soft (minor) and hard (major) page faults.
% # Context switches: Voluntary and involuntary context switching events.
% \end{lstlisting}

\begin{lstlisting}[language=C, caption={In-enclave system-level instrumentation using \texttt{getrusage()}}, label={lst:getrusage}]
struct rusage usage;
getrusage(RUSAGE_SELF, &usage);

// CPU time: Total time spent in user mode and kernel mode.
printf("!User CPU time:     %ld.%06lds\n", usage.ru_utime.tv_sec, usage.ru_utime.tv_usec);
printf("!System CPU time:   %ld.%06lds\n", usage.ru_stime.tv_sec, usage.ru_stime.tv_usec);

// Peak RSS: Maximum resident memory usage during execution.
printf("!Max memory (RSS):  %ld KB\n", usage.ru_maxrss);

// Page faults: Number of soft (minor) and hard (major) page faults.
printf("!Minor page faults: %ld\n", usage.ru_minflt);
printf("!Major page faults: %ld\n", usage.ru_majflt);

// Context switches: Voluntary and involuntary context switching events.
printf("!Voluntary context switches:   %ld\n", usage.ru_nvcsw);
printf("!Involuntary context switches: %ld\n", usage.ru_nivcsw);
\end{lstlisting}

These runtime indicators were critical in diagnosing performance behaviors that could not be attributed to CPU throughput alone. For instance, elevated context switch counts in parallel configurations were often correlated with increased enclave transition overhead or CPU oversubscription, especially in memory-constrained environments.

Cache-related metrics were obtained using the QEMU TCG plugin cache model described earlier. This plugin logs all instruction and data cache accesses and cache misses at the system level. While the plugin lacks enclave-specific visibility and does not model hardware-specific cache timing, it provides a consistent and useful proxy for analyzing trends in memory access behavior. Cache metrics were aligned with benchmark execution phases to identify correlations between cache activity and observed performance shifts.

To compute the enclave-specific performance overhead, we employed a normalized overhead metric defined as the relative increase in execution time between enclave and native modes:

\[
\text{Overhead (\%)} = \left( \frac{\text{Enclave Time} - \text{Native Time}}{\text{Native Time}} \right) \times 100
\]

This percentage-based formulation allows for direct cross-comparison of overheads across workloads with differing baseline durations. It also facilitates trend identification across system configurations, particularly when assessing scaling behaviors.

For reproducibility, all benchmarking scripts, system configurations, QEMU command-line options, Keystone runtime parameters, and cache plugin settings were version-controlled using Git. Raw benchmark outputs, kernel logs, and cache plugin trace files were archived and documented. A complete snapshot of the experimental infrastructure, along with scripts to regenerate results, is provided in Appendix~\ref{appendix:cache-plugin}. This archival effort ensures that all reported results can be independently verified and reproduced by future researchers or evaluators.