### 2025-08-31 draft_presentation.tex
added slides

\documentclass[aspectratio=169]{beamer}
\usetheme{metropolis}

\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{graphicx}
\usepackage{booktabs}
\usepackage{fontawesome5} % for icons
\usepackage{pifont}       % for checkmarks
\usepackage{xcolor}       % for color definitions
\usepackage{colortbl}     % for \rowcolor, \rowcolors, \cellcolor in tables
\usepackage{tikz}         % for drawing graphics
\usetikzlibrary{patterns,positioning} % enable patterns and positioning for tikz
\usepackage{pgfplots} 
%\pgfplotsset{compat=1.17} % set PGFPlots compatibility

% Colours
\definecolor{keystoneblue}{RGB}{0,77,128}
\definecolor{warmaccent}{RGB}{255,140,0}
\setbeamercolor{structure}{fg=keystoneblue}
\setbeamertemplate{navigation symbols}{}
\setbeamertemplate{footline}[frame number]

% Styling tweaks
\metroset{block=fill,progressbar=frametitle,numbering=fraction}
\renewcommand{\baselinestretch}{1.05}
\setbeamertemplate{frametitle}[default][center]

% Title and Author
\title[Evaluating RISC-V Enclaves]{\textbf{Evaluating RISC-V Enclaves:}\\ Security Without the Slowdown?}
\author{\textbf{Basil Ugbomoiko}}
\institute[UzL]{Universität zu Lübeck \\ Institut für Technische Informatik}
\date{Master’s Thesis Defense — \today}
% Add logo (adjust path as needed)
\titlegraphic{\hfill\includegraphics[height=1.5cm]{uzl-logo.pdf}}
\begin{document}

% Title Page with University Logo (if available)
\begin{frame}
  \titlepage
  \vfill
  \centering\small\textit{In the age of untrusted infrastructure, can security keep up with performance?}
\end{frame}

% Opening Question Frame (Increased vertical space)
\begin{frame}{Opening Question}
\centering
\Huge\bfseries
Can we \textcolor{keystoneblue}{trust} computation on machines we \textcolor{warmaccent}{don't control}? \\[1em]
\large In cloud and edge scenarios, this question is no longer optional — it's urgent.
\end{frame}

\begin{frame}{Problem Statement: Trends}
\begin{itemize}
    \item[\faCloud] Growing reliance on \textcolor{warmaccent}{cloud} and \textcolor{warmaccent}{edge} computing.
    \item[\faUsers] Shared, multi-tenant infrastructure is the norm.
\end{itemize}
\end{frame}

\begin{frame}{Problem Statement: Challenge}
\begin{itemize}
    \item[\faExclamationTriangle] Traditional security assumes a trustworthy OS — \alert{assumption broken}.
    \item[\faLock] Need isolation of critical code/data \textbf{even if the OS is compromised}.
\end{itemize}
\vfill
%\scriptsize\textit{This sets the stage for Trusted Execution Environments (TEEs).}
\end{frame}

% Background on TEEs (Added icons for clarity)
\begin{frame}{Background: TEEs in a Nutshell}
\begin{block}{Definition}
Hardware-based secure “zones” for executing sensitive code and storing sensitive data.
\end{block}
Examples include:
\begin{itemize}
    \item Intel SGX \faMicrochip\ \textit{(Intel's enclave solution for secure execution)} 
    \item ARM TrustZone \faLock\ \textit{(Security for mobile devices)}
    \item AMD SEV \faServer\ \textit{(Virtual machine encryption for server environments)}
    \item Keystone \faKey\ \textit{(Modular, open-source for RISC-V)}
\end{itemize}
\end{frame}

% Research Gap (Emphasized visual structure)
\begin{frame}{Research Gap}
\centering
\Large\bfseries
No comprehensive, benchmarked assessment of \\
\textbf{Keystone} performance overhead with \\
\textit{actionable tuning guidance.}
\end{frame}

% % Research Objectives (Visual appeal with bullet style)
% \begin{frame}{Research Objectives}
% \begin{enumerate}
%     \item Characterize Keystone’s \textbf{architecture} and \textbf{security model}.
%     \item Quantify performance across \textcolor{warmaccent}{varied workloads}.
%     \item Recommend \textbf{best practices} for optimal configuration.
% \end{enumerate}
% \end{frame}

% What is RISC-V? Slide
\begin{frame}{What is RISC-V?}
\begin{itemize}
    \item Open-source, modular processor architecture.
    \item Highly customizable for specialized hardware (ideal for enclaves).
    \item Scalable from embedded systems to high-performance computing.
    \item Growing adoption in academia and industry (e.g., SiFive, Western Digital).
\end{itemize}
\begin{block}{Key Concept}
\textit{RISC-V’s open-source nature enables flexible, secure processor designs for enclave-based systems.}
\end{block}
\end{frame}

% Keystone 
\begin{frame}{Why Keystone?}
\begin{itemize}
    \item Modular, open-source TEE for RISC-V.
    \item Uses Physical Memory Protection (PMP) and minimal Security Monitor for isolation.
    \item Low Trusted Computing Base (TCB) for security and flexibility.
    \item Optimized for secure computation in untrusted environments.
\end{itemize}
\begin{block}{Key Concept}
\textit{Low TCB reduces attack surfaces while enabling high flexibility and performance in enclave-based systems.}
\end{block}
\end{frame}

% Keystone Framework Architecture
\begin{frame}{Keystone Framework Architecture}
\begin{block}{Overview}
    Keystone is designed to be modular, with clear isolation between enclaves and the OS.
\end{block}

%\resizebox{\linewidth}{!}{%
\begin{figure}[htbp]
\centering
\includegraphics[width=\linewidth]{/home/individual-01/Documents/keystone-thesis/figures/keystone_overview.png}
%\caption{Keystone architecture showing components such as the Security Monitor, enclave runtime, and the privilege hierarchy.}
\label{fig:keystone_grouped}
\end{figure}
\end{frame}

% Keystone vs Other TEEs
\begin{frame}{Keystone vs. Other TEEs}
\begin{figure}[htbp]
\centering
\caption{Feature comparison of major TEEs.}
\label{fig:tee-features}
\renewcommand{\arraystretch}{1.3} % more row height
\setlength{\tabcolsep}{5pt}       % more column padding
\rowcolors{2}{gray!5}{white}      % alternating row colors

\begin{tabular}{@{}lcccc@{}}
\toprule
\rowcolor{gray!20}
\textbf{Feature} & \textbf{Keystone} & \textbf{SGX} & \textbf{TrustZone} & \textbf{SEV} \\
\midrule
Open Source            & \cellcolor{green!15}Yes & \cellcolor{red!15}No & \cellcolor{red!15}No & \cellcolor{red!15}No \\
Dynamic Enclaves       & \cellcolor{green!15}Yes & \cellcolor{red!15}No & \cellcolor{yellow!15}Limited & \cellcolor{red!15}No \\
Hardware Dependency    & \cellcolor{green!15}RISC-V & \cellcolor{red!15}Intel & \cellcolor{red!15}ARM & \cellcolor{red!15}AMD \\
TCB Size               & \cellcolor{green!15}Low & \cellcolor{yellow!15}Medium & \cellcolor{red!15}High & \cellcolor{yellow!15}Medium \\
Flexible Configurations& \cellcolor{green!15}High & \cellcolor{red!15}Low & \cellcolor{red!15}Low & \cellcolor{yellow!15}Medium \\
\bottomrule
\end{tabular}
\end{figure}

\end{frame}

\begin{frame}{System Design and Methodology}
    \small
    \begin{itemize}
        \item \textbf{Objective:} Evaluate the performance impact of Keystone's enclave isolation mechanisms on embedded/system-level workloads. \pause
        \item \textbf{Platform:} Emulated RISC-V architecture (RV64GC) using QEMU. \pause
        \item \textbf{Focus:} Quantify computational overhead when using Keystone enclaves vs. traditional non-isolated execution. \pause
    \end{itemize}
    \vspace{0.5em}
    \uncover<4->{\textbf{Platform Configuration: QEMU-based emulation ensures controlled environment for testing.}}
\end{frame}

\begin{frame}{QEMU-Based Virtualized Environment}
    \small
    \begin{itemize}
        \item QEMU emulates RISC-V architecture for repeatability and control. \pause
        \item Ensures consistent test conditions by eliminating hardware-induced variations (thermal throttling, cache behavior). \pause
        \item Configured for RV64GC architecture, as required by Keystone. \pause
    \end{itemize}
    \vspace{0.5em}
    \uncover<4->{\textbf{Why QEMU?} Provides a controlled, consistent environment for repeatability.}
\end{frame}

\begin{frame}{Benchmarking Workflow}
    \small
    \begin{itemize}
        \item \textbf{Define benchmark set} and configure system setup. \pause
        \item \textbf{Dual-mode execution:} Run benchmarks in native mode (without enclave) and enclave mode (with isolation). \pause
        \item \textbf{Metrics collection:} Gather data on execution time, throughput, and overhead. \pause
        \item \textbf{Analysis:} Compare performance across modes and assess overheads. \pause
    \end{itemize}
    \vspace{0.5em}
    %\uncover<5->{\textbf{Visual Workflow:} As shown in Figure \ref{fig:benchmarking-swimlane}, the process starts with benchmarking selection and proceeds through dual-mode execution.}
    \begin{figure}[H]
    \centering
\begin{figure}[htbp]
\centering
\resizebox{\linewidth}{!}{%
\begin{tikzpicture}[
    >=stealth,
    phase/.style={
        font=\bfseries\small,
        text=white,
        minimum height=0.8cm,
        rounded corners,
        inner xsep=6pt,
        text centered
    },
    stepnode/.style={
        draw=black!70,
        thick,
        rounded corners,
        align=center,
        font=\scriptsize,
        minimum height=1.4cm,
        text width=2.8cm,
        inner sep=6pt
    },
    arrow/.style={->, thick, shorten >=2pt, shorten <=2pt}
]

% Top row (Setup & Config) - blue shades
\node[stepnode, fill=blue!15] (s1) at (0,3) {\faTasks\\Define benchmarks};
\node[stepnode, fill=blue!20, right=1.8cm of s1] (s2) {\faMicrochip\\Select hardware};

% Bottom row (Execution & Analysis) - green shades
\node[stepnode, fill=green!15, right=1.8cm of s2] (s3) {\faPlay\\Run native mode};
\node[stepnode, fill=green!20, right=1.8cm of s3] (s4) {\faLock\\Run enclave mode};
\node[stepnode, fill=green!25] (s5) at (0,1) {\faChartBar\\Collect metrics};
\node[stepnode, fill=green!30, right=1.8cm of s5] (s6) {\faSearch\\Analyze overheads};
\node[stepnode, fill=green!40, right=1.8cm of s6] (s7) {\faArchive\\Archive results};

% Arrows top row
\draw[arrow] (s1) -- (s2);
\draw[arrow] (s2) -- (s3);
\draw[arrow] (s3) -- (s4);

% Arrow from top to bottom
\draw[arrow] (s4.south) .. controls +(0,-0.8) and +(0,0.8) .. (s5.north);

% Arrows bottom row
\draw[arrow] (s5) -- (s6);
\draw[arrow] (s6) -- (s7);

% Color legend
\node[draw=black!50, fill=white, rounded corners, align=left, font=\scriptsize, anchor=north west] 
    at ([xshift=1cm,yshift=-0.5cm]s7.south east) (legend) {
    \begin{tikzpicture}[baseline=(current bounding box.center)]
        \node[fill=blue!20, minimum width=0.5cm, minimum height=0.5cm] (b) {};
        \node[right=0.3cm of b, anchor=west] {Setup \& Config};
        \node[fill=green!30, minimum width=0.5cm, minimum height=0.5cm, below=0.2cm of b] (g) {};
        \node[right=0.3cm of g, anchor=west] {Execution \& Analysis};
    \end{tikzpicture}
};

\end{tikzpicture}
}
\caption{Benchmarking and data collection workflow}
\end{figure}
    \caption{Benchmarking and data collection workflow}
    \label{fig:benchmarking-swimlane}
    \end{figure}
\end{frame}

\begin{frame}{Benchmarking Tools}
    \small
    \begin{itemize}
        \item \textbf{Dhrystone:} Measures CPU performance (DMIPS). \pause
        \item \textbf{CoreMark:} Focuses on integer performance, used in embedded systems benchmarking. \pause
        \item \textbf{Two configurations:}
        \begin{itemize}
            \item \textbf{Native Execution:} Benchmark runs as a normal process. \pause
            \item \textbf{Enclave Execution:} Benchmark runs within a Keystone enclave with memory isolation. \pause
        \end{itemize}
    \end{itemize}
    \vspace{0.5em}
    \uncover<5->{\textbf{Key Focus:} These tools provide insights into both CPU-bound and system-level performance.}
\end{frame}

%Results
\section{Synthesis of Key Performance Trends and Hardware Bottlenecks}
\label{sec:synthesis}

\begin{frame}{Enclave Overhead and System Behavior}
\small
\begin{block}{Performance Overhead}
\begin{itemize}
    \item \textbf{Enclave execution} introduces overhead: 10\% to 15\%.
    \item \textbf{Compute-bound workloads} (CoreMark, Dhrystone) see minimal performance loss.
    \item \textbf{Memory-bound workloads} (Kyber) experience severe slowdowns due to added security checks and memory isolation.
\end{itemize}
\end{block}

\begin{block}{Overhead Breakdown}
\begin{itemize}
    \item Memory isolation via PMP adds fixed latency per memory access.
    \item Context switches between host and enclave domains are costly.
\end{itemize}
\end{block}

\centering
\textbf{Key Insight:} Overhead varies based on workload characteristics and memory behavior.
\end{frame}

\begin{frame}{Impact of Scalability Limitations}
\small
\begin{block}{CPU Scaling Limits}
\begin{itemize}
    \item Increasing CPU cores \textbf{benefits multi-threaded tasks}, but gains plateau due to context-switch overhead and enclave thread management.
    \item For example, \texttt{CoreMark} shows marginal improvements beyond 4 cores.
\end{itemize}
\end{block}

\begin{block}{Memory Scaling Constraints}
\begin{itemize}
    \item Increased memory capacity \textbf{improves performance} by reducing page faults.
    \item \textbf{Fragmentation} and \textbf{PMP entry limits} cap the performance improvement.
\end{itemize}
\end{block}

\centering
\textbf{Key Insight:} Hardware resources alone cannot fully predict enclave performance due to system resource allocation limits.
\end{frame}

\begin{frame}{PMP Entry Exhaustion and Multi-Enclave Scalability}
\small
\begin{block}{PMP Resource Limitations}
\begin{itemize}
    \item \textbf{Maximum PMP entries}: 16.
    \item Exhausted PMP entries \textbf{prevent new enclave creation}, causing errors even with abundant system resources.
    \item Failure due to PMP exhaustion: \texttt{SBI error codes}.
\end{itemize}
\end{block}

\begin{block}{Scalability Bottleneck}
\begin{itemize}
    \item In \textbf{multi-tenant environments}, PMP exhaustion limits the number of concurrent enclaves.
    \item This creates scalability issues, especially for high-concurrency workloads.
\end{itemize}
\end{block}

\centering
\textbf{Key Insight:} PMP limitations represent a significant scalability bottleneck for RISC-V enclave-based systems.
\end{frame}

\begin{frame}{Performance Trends and Hardware Bottlenecks}
\centering
\begin{figure}[htbp]
\centering
\begin{tikzpicture}
\begin{axis}[
    ybar,
    bar width=14pt,
    width=10cm,
    height=6cm,
    ymin=0, ymax=120,
    ylabel={Relative performance (\%)},
    symbolic x coords={CoreMark,Dhrystone,Kyber},
    xtick=data,
    ymajorgrids,
    grid style=dashed,
    nodes near coords,
    every node near coord/.append style={font=\footnotesize, /pgf/number format/fixed},
    legend style={at={(0.5,-0.15)}, anchor=north, legend columns=-1}
]

% Enclave: blue, rounded top, transparency
\addplot[
    fill={rgb,255:red,0; green,114; blue,178},
    fill opacity=0.8,
    draw=black!60,
    rounded corners=2pt
] coordinates {(CoreMark,95) (Dhrystone,62) (Kyber,37)};

% Native: orange, rounded top, transparency
\addplot[
    fill={rgb,255:red,230; green,159; blue,0},
    fill opacity=0.8,
    draw=black!60,
    rounded corners=2pt
] coordinates {(CoreMark,100) (Dhrystone,100) (Kyber,100)};

\legend{Enclave,Native}
\end{axis}
\end{tikzpicture}
\caption{Relative performance of enclave vs native execution.}
\end{figure}
\end{frame}

\begin{frame}{Key Takeaways from Performance Trends}
\small
\begin{block}{Main Findings}
\begin{itemize}
    \item Enclave performance suffers primarily in \textbf{memory-bound workloads} due to security checks and context switching.
    \item \textbf{PMP entry limitations} restrict scalability in multi-enclave setups.
    \item Increasing hardware resources does not linearly translate to performance gains due to system-level overheads.
\end{itemize}
\end{block}

\centering
\textbf{Key Insight:} Optimizing both hardware resources and system-level configurations is essential for scalable and efficient enclave execution.
\end{frame}

\begin{frame}{Why Misalignment Bloats PMP Usage}
\small
\vfill

% Problem
\only<1>{
\begin{block}{Problem}
Physical Memory Protection (PMP) entries must cover power‑of‑two sized, naturally aligned regions (NAPOT).
\end{block}
}

% Cause
\only<2>{
\begin{block}{Problem}
Physical Memory Protection (PMP) entries must cover power‑of‑two sized, naturally aligned regions (NAPOT).
\end{block}
\vspace{0.5em}
\begin{block}{Cause}
When a memory region is \textbf{misaligned}:
\begin{itemize}
  \item NAPOT encoding forces padding to the next power‑of‑two boundary.
  \item Padding areas consume PMP coverage even if they are not actual data.
\end{itemize}
\end{block}
% \draw[->, thick] (0,-0.2) -- (0,-0.8); % optional arrow from Problem → Cause
}

% Effect
\only<3>{
\begin{block}{Cause}
When a memory region is \textbf{misaligned}:
\begin{itemize}
  \item NAPOT encoding forces padding to the next power‑of‑two boundary.
  \item Padding areas consume PMP coverage even if they are not actual data.
\end{itemize}
\end{block}
\vspace{0.5em}
\begin{block}{Effect}
\begin{itemize}
  \item A single logical region may require \textbf{multiple PMP entries}.
  \item Fragmentation increases as more regions are misaligned.
\end{itemize}
\end{block}
% \draw[->, thick] (0,-0.2) -- (0,-0.8); % optional arrow from Cause → Effect
}

% Impact
\only<4>{
\begin{block}{Effect}
\begin{itemize}
  \item A single logical region may require \textbf{multiple PMP entries}.
  \item Fragmentation increases as more regions are misaligned.
\end{itemize}
\end{block}
\vspace{0.5em}
\begin{block}{Impact}
\begin{itemize}
  \item Higher PMP entry count $\Rightarrow$ fewer entries left for other regions.
  \item Limits scalability and flexibility of memory protection.
  \item Wastes address space and complicates security policy.
\end{itemize}
\end{block}
}

% Key takeaway
\only<5>{
\centering
\begin{figure}[htbp]
\centering
\begin{tikzpicture}[scale=1, transform shape,font=\small]

% Memory axis
\draw[thick,->] (0,0) -- (10,0) node[anchor=west]{Memory address};
\foreach \x/\label in {0/0\,KB,1.25/128\,KB,4.25/192\,KB,8/300\,KB}
    \draw (\x,0.1) -- (\x,-0.1) node[anchor=north]{\label};

% === Region 1: contiguous, 1 PMP entry ===
\node[anchor=south] at (1.375,2.1) {\scriptsize 128\,KB aligned};
%\draw[pattern=north east lines,pattern color=black] (0.25,0) rectangle (0.5,1);
\draw[fill=keystoneblue!60] (0.5,0) rectangle (2.2,1);
%\draw[pattern=north east lines,pattern color=black] (2.2,0) rectangle (2.5,1);
\draw[fill=keystoneblue!60,opacity=0.4] (0.5,1) rectangle (2.2,2) node[pos=.5] {\scriptsize 1 entry};

% === Region 2: fragmented, 3 PMP entries ===
\node[anchor=south] at (4.3,4.1) {\scriptsize 192\,KB misaligned};
\draw[pattern=north east lines,pattern color=black] (3.5,0) rectangle (3.8,1);
\draw[fill=warmaccent!60] (3.8,0) rectangle (4.6,1);
\draw[pattern=north east lines,pattern color=black] (4.6,0) rectangle (5.1,1);
\draw[fill=warmaccent!60,opacity=0.4] (3.5,1) rectangle (5.1,4) node[pos=.5] {\scriptsize 3 entries};

% === Region 3: fragmented, 3 PMP entries ===
\node[anchor=south] at (8.0,4.1) {\scriptsize 300\,KB misaligned};
\draw[pattern=north east lines,pattern color=black] (6.7,0) rectangle (7.0,1);
\draw[fill=warmaccent!60] (7.0,0) rectangle (9.1,1);
\draw[pattern=north east lines,pattern color=black] (9.1,0) rectangle (9.3,1);
\draw[fill=warmaccent!60,opacity=0.4] (6.7,1) rectangle (9.3,4) node[pos=.5] {\scriptsize 3 entries};

% Arrows showing cause-effect
%\draw[->,thick] (2.2,0.5) .. controls (2.8,1.5) .. (3.4,2) node[midway,above,sloped]{\scriptsize Padding $\Rightarrow$ More entries};
\draw[->,thick] (5.1,0.5) .. controls (5.4,2) .. (6.6,2.5) node[midway,above,sloped]{\scriptsize Padding $\Rightarrow$ More entries};

% Legend
\begin{scope}[xshift=0cm,yshift=-1.5cm]
\draw[fill=keystoneblue!60] (0,0) rectangle (0.4,0.4) node[right=0.5cm] {Actual region (contiguous)};
\draw[fill=warmaccent!60] (5,0) rectangle (5.4,0.4) node[right=0.5cm] {Actual region (fragmented)};
\draw[pattern=north east lines,pattern color=black] (0,-0.6) rectangle (0.4,-0.2) node[right=0.5cm] {Wasted padding};
\draw[fill=gray!50,opacity=0.4] (5,-0.6) rectangle (5.4,-0.2) node[right=0.5cm] {PMP entry count bar};
\end{scope}

\end{tikzpicture}

\vspace{0.4em}
\scriptsize
\textbf{Takeaway:} Align regions to NAPOT boundaries to minimize padding and preserve PMP entries.
\end{figure}
}

\vfill
\end{frame}

% Optimizations for Keystone Enclaves
\begin{frame}{Optimizations for Keystone Enclaves}
\begin{itemize}
    \item Use contiguous, NAPOT-aligned memory to minimize PMP usage.
    \item Consolidate workloads and limit active enclaves to reduce contention.
    \item Offload non-sensitive tasks to the untrusted domain.
\end{itemize}
\end{frame}

% Real-World Applications of Keystone Enclaves
\begin{frame}{Real-World Applications of Keystone Enclaves}
\begin{itemize}
    \item \textbf{Secure Cloud Computing:} Protects sensitive data in untrusted environments, ensuring confidentiality in multi-tenant clouds.
    \item \textbf{Confidential Data Processing:} Enables secure computations on private data, ideal for financial and healthcare industries.
\end{itemize}
\vspace{2em}
\resizebox{\textwidth}{!}{ % Scale the entire diagram
\begin{tikzpicture}[
    node distance=1.5cm, % Reduce space between nodes
    every node/.style={font=\small, align=center},
    box/.style={draw, rounded corners, thick, fill=blue!15, minimum width=2.8cm, minimum height=1.2cm},
    data/.style={draw, thick, fill=orange!20, rounded corners, minimum width=2.5cm, minimum height=1cm}
]

% Cloud
\node[box, fill=blue!20] (cloud) {\includegraphics[scale=0.05]{example-image} \\ \textbf{Cloud Server} \\ (Untrusted Environment)};

% Lock / Enclave
\node[box, fill=green!20, right=3cm of cloud] (enclave) {\faLock \\ \textbf{Keystone Enclave} \\ Secure Execution};

% Data source
\node[data, left=3cm of cloud] (client) {\faUser \\ \textbf{Client Data}};

% Arrows
\draw[->, thick] (client) -- node[above]{Encrypted Data} (cloud);
\draw[->, thick] (cloud) -- node[above]{Secure Transfer} (enclave);
\draw[->, thick] (enclave) -- ++(0,-2) node[midway,right]{Results (Encrypted)} -| (client.south);

% Optional note
\node[below=1.2cm of cloud, text width=8cm, align=center] 
    {\textit{Sensitive data is encrypted before leaving the client, processed inside the enclave, and never exposed in plaintext to the cloud provider.}};

\end{tikzpicture}
}
\end{frame}

% Future Work and Improvements
\begin{frame}{Future Work and Improvements}
\begin{itemize}
    \item Performance optimizations with hardware extensions.
    \item Multi-enclave systems for more complex use cases.
    %\item Research on side-channel attacks and improved security measures.
    \item Development of better developer tools for enclave management.
\end{itemize}
\end{frame}

% Summary of Findings
\begin{frame}{Key Takeaways}
\begin{itemize}
    \item Keystone offers strong security with minimal overhead for CPU-bound workloads.
    \item Significant overhead for memory bound tasks can be reduced with optimizations.
    \item Customizable settings allow for optimization based on workload type.
\end{itemize}
\end{frame}

% Conclusion
\begin{frame}{Conclusion}
\centering
\Large Keystone can deliver strong security with minimal performance cost — \\
\textit{if workloads are chosen and designed wisely.}
\end{frame}

% Thank You Slide with visual focus
\begin{frame}
\centering
\Huge\textbf{Thank You}\\
\large Questions? \faQuestionCircle
\vfill
\small Universität zu Lübeck — Master’s Thesis Defense
\end{frame}
\end{document}